Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector

### 2018-08-03 documentation/paper.tex
Fixed statistics; added section on image CNN

Once the audio was cropped properly, $DeepConv(\omega)$ achieved a maximum of 77\% accuracy on validation data. Using the same network architecture (except for the input layer), $DeepConvWithPos(\omega, \chi)$ managed 85\%. The maximum obtained on a grid search with position correction was 91\%. While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume. That said, neither of these accuracy statistics are particularly good, which does not bode well for this input format as a whole.

@@

A multi-layer perceptron (a neural network composed of fully connected layers) was applied to analysis of audio data in the frequency domain. Preprocessing by applying a Fourier transform makes it an easier task for the neural network to accomplish in some ways. If the frequency distribution and overall volume are indeed the most important factors for discrimination, the network can gather them straight from the Fourier transform $\beta_{N}$ rather than having to analyze $\omega$ to extract this information.

@@

>>>>

Once the audio was cropped properly, $DeepConv(\omega)$ achieved a maximum of 77\% accuracy on validation data. Using the same network architecture (except for the input layer), $DeepConvWithPos(\omega, \chi)$ managed 85\%. The maximum obtained on a grid search with position correction was 91\%. While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume. That said, neither of these accuracy statistics are particularly good (graphs of 85\% accuracy often appear effectively random), which does not bode well for this input format as a whole.

@@

A multi-layer perceptron (a neural network composed of dense layers) was applied to analysis of audio data in the frequency domain. Preprocessing by applying a Fourier transform makes it an easier task for the neural network to accomplish in some ways. If the frequency distribution and overall volume are indeed the most important factors for discrimination, the network can gather them straight from the Fourier transform $\beta_{N}$ rather than having to analyze $\omega$ to extract this information.

@@

\subsection{Convolutional Neural Network for Image Window Analysis}

It is an open question whether or not there is any information in the image data $\iota$ that could be used to distinguish between particle classes. Relative to the extremely short period of time in which the bubble forms (on the scale of nanoseconds), the framerate of the camera (340Hz) is extremely slow. While the very early stages of bubble formation (when the sound is produced) are known to differ dependong on whether the bubble was created by a nuclear recoil or an alpha particle, it is unknown whether any visually apparent differences persist when the bubble is visible.

In an effort to resolve this, a 2-dimensional convolutional neural network was applied to the task of discriminating based on $\iota$. The network architecture consists of a moderate number of convolutional layers (on the scale of 9) and 3 dense layers at the end. L2 regularization was used on all layers, and dropout was additionally used on the dense layers. No position input is used (since there are no microphones that would require position correction), so the basic configuration is $ImageConv(\iota)$.

Throughout a variety of different network architectures, the best validation accuracy obtained was 85\%. (The corresponding training accuracy of 100\% indicates that severe overfitting is taking place.) It is likely fitting on some form of noise within $\iota$, since the network's predictions seem almost random when represented graphically. The fact that throughout many trials (including a grid search), no good performance on validation data was observed, provides significant evidence that images provide insufficient information for effective discrimination.
